---
title: "Functions in Python and R"
author: "Sharon C. Makunura"
date: "2022-08-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this article I describe how to define your own functions in both Python and R.  I find that the main difference in using functions for either of these programming languages lies mainly with the output, as I will demonstrate. 

In case you didn't know, functions are reusable code that take input and produce output, are reusable and can be called multiple times.

There are two main section to creating functions: i. defining the function and its arguments, and ii. specifying the actions of the function.  In this example the aim is to create a function that will double, half and square a number.

## 1. Define a simple Python function

A Python function follows the following format:

**def** function_name(arguments):


|         do something


Working with our example to double, half and square a number the function looks looks as follows. 

```{python define_function, echo=TRUE }
def manipulate(x):
  print("double is ",(x*2))
  print("half is ",(x/2))
  print("squared is ",(x*x))
```

The code above defines the function.  To use the function, we need to call the function as shown below:

```{python}
manipulate(5)
```



## 2. Define a simple R function  

The format of a function in R is as follows:


function_name <- **function** (arguments){


|         do something
|         **return** (output)

}

Again, working with the same example as above will give us the following function.  
```{r}
manipulateR <- function(x){
  double <-x*2
  halved <-x/2
  squared <-x*x
  
  results<-list(double, halved, squared)
  names(results)<- c("Doubled", "Halved", "Squared")
  return(results)
}
```
All functions need to be called to be able to use them.  

```{r}
manipulateR(5)
```
## 3. The difference 


For the most part function in Python and R are constructed and used in a very similar manner. I used the print function for my Python output and a data frame for my R function.  The latter is necessitated by a unique characteristic of R - it can only return a single output.  This is unlike Python which supports multiple outputs as illustrated in the following example.

First I will redefine the function so that instead of printing, each result is saved as an individual variable.  

```{python new_function, echo = FALSE}
def manipulate(x):
  d =x*2
  h=x/2
  s=x*x
  return(d,h,s)
```
When calling the function, the output is a list.

```{python}
manipulate(5)

```
The advantage of this structure of output is that each value can be independently assigned to variables as demonstrated below:

```{python}
x,y,z = manipulate(5)
print("double is ",x)
print("half is ",y)
print("squared is ",z)
```
In R, only a single object can be the output from the function, for example a data frame as shown above.  Take a look at [this R project](https://github.com/SharonMakunura/Google-Data-Analytics-Capstone-Project/blob/39936c05faa52474df5e9e8921273dd22500b066/cyclistic-analysis-script.R) and [this Python project](https://github.com/SharonMakunura/WhatsApp-Data-Analysis-With-Python/blob/a4811e4e1023c3efb4c1720af382776ff63a4b0f/3.%20Analyzing%20Whatsapp%20Data.ipynb) to see how I implemented customized functions and different outputs.




